Automated Lensing Learner - I: An automated Strong Lensing Identification Pipeline

### 2017-03-21 ms.tex
Went through methodology additions and highlighted in purple re-wording that Matt needs to look at

% Image identification works: visual, ``robot'' identification -
% ArcFinder, RingFinder..., and citizen science (space warps).  Many
% of these rely on some explicit aspect of the image's morphology
% (give example), filtering, ... ....  Specifics of the methodology
% how the instrument affects the strong lensing signal, and the
% objects of interest.
Over the last decade, infrastructure for both large scale visual and
automated image classification emerged. By nature, the human eye is
one of the best discriminators for image classification.  Visual arc
identification has been effective through the use of citizen science
platforms.  {\em SpaceWarps} is an example of citizen science based
image classification of strong lensing systems in Canada France Hawaii
Telescope Science (CFHTLS) telescope observations
\citep{marshall_etal16,more_etal16}.  These platforms are quite
successful for a dataset like CFHTLS, classifying \todo{XX in XX
  time}\citep{}.  However, future datasets like Euclid
\citep{oguriandmarshall_10} and LSST \cite{} will be \todo{XX} times
this size, and {\em Euclid} is estimated to find tens of thousands
galaxy-scale strong lenses \citep{pawase_etal12}.  The volume of
upcoming data challenges the scalability of a citizen science
approach.

@@

% Convolutional and residual
The most advanced use of neural networks for strong lensing
classification operate directly in image space.
\citep{petrillo_etal17} used mock Kilo Degree Survey data to train
convolutional neural networks, with a training set size of six million
images. \citet{lanusse_etal17} used state-of-the-art residual learning
to also work directly in the image space with minimal image
pre-processing.  A major strength of the \citet{lanusse_etal17}
implementation is that residual learning has been found to be easier
to train in deep neural networks, and can outperform implementations
of convolutional neural networks.

%  Relevance of our work
While there has been a recent surge in the use of neural networks
applied to astronomy image classification problems, \wording{these
  implementations are not readily scalable, nor are the necessary
  computational resources and expertise easily accessible to the
  entire scientific community.}  We present a supervised
classification pipeline built with open source tools, tested on mock
HST and LSST data, which will also be made publicly available.  Our
goal is to centralize an open source pipeline that is general enough
for the user to select or add appropriate image processing techniques
to augment the features of the lensed image, train and test data with
different machine learning techniques, and to select the sequence of
methods that provides optimal model performance.  Results are
reproducible on personal computers, as both the pipeline and the data
will be publicly distributed.

% Summary of paper 
This first paper serves as a case study for our pipeline with one
feature extraction and classical machine learning method,
respectively, the {\em histogram of oriented gradients} (HOG) and {\em
  logistic regression} (LR).  HOG is a fast feature extraction
procedure that quantifies edges in images, commonly used to identify
humans in security software.  LR is a linear classifier, making its
scalability relatively straightforward with existing open source
tools.  This framework is easily extendable to other feature methods
and machine learning algorithms for supervised classification.

@@

%-------------------------------------------------%
\section{Methodology}
\label{sec:methods}
%-------------------------------------------------%
\subsection{Mock Observations}
%-------------------------------------------------%
\subsection{Mock Images}
%-------------------------------------------------%
% Summary of HST and LSST images
We use mock {\em Hubble Space Telescope}\footnote{\todo{hst...org}}
(HST) and {\em Large Synoptic Sky
  Telescope}\footnote{\todo{lsst....org}} (LSST) images generated with
PICS (Pipeline for Images of Cosmological Strong lensing)
\citep{li_etal16} and LensPOP \citep{collett_15} to train and test our
model.  Along the line of sight galaxies and foreground stars.  Half
of the images used are strong lensing systems.  Lensed galaxy images
are ray-traced images of actual galaxies extracted from deep Hubble
Space Telescope observations.  The final mock images are ``observed''
with a realistic point spread function corresponding to modeled
detector artifacts for bright stars.  We refer to \cite{li_etal16} for
further details, but outline the process here.

We generate 10,000 lens containing mock observations and 10,000
non-lens containing mock observations to run our parameter search, and
keep a hold-out test set of 1,000 lens containing mock observations
and 1,000 non-lens containing mock observations on which we evaluate
the final trained model.

Lens containing mock observations include the lens galaxy, lensed
images of the source galaxy, and galaxies along the line of sight.
Non-lens containing mock observations include all but a lensed source
image.  We convolve each of the $2\times 10,000$ train/test images and
$2\times 1,000$ hold-out images to produce three separate sets of mock
``observations'' with equal numbers of lens and non-lens containing
systems: (1) HST-like observations, (2) LSST-like observations over
the span of one year, and (3) LSST-like observations over the span of
ten years.  We respectively label these observations as HST, LSST1,
and LSST10.
%-------------------------------------------------%
\subsubsection{Modeling Images of the Source Galaxies}
%-------------------------------------------------%
To produce images of the simulated lensed source galaxies, we first
set mass models of the lens galaxy to be the Singular Isothermal
Ellipsoid (SIE):
%-------------------------------------------------%
\begin{equation}\label{eqn:sie}
\kappa = \frac{\theta_{\rm E}}{2}\frac{1}{\sqrt{x_1^2/q+x_2^2 q}},
\end{equation}
%-------------------------------------------------%
where $\theta_{\rm E}$ is the Einstein radius and $q$ is the axis
\wording{ratio between semi-major and semi-minor axes $x_1$ and
  $x_2$}. We can calculate the Einstein radius from the redshift of
the lens, the redshift of the source, and the velocity dispersion of
the lens galaxy.
%-------------------------------------------------%
\begin{equation}\label{eqn:einsteinradius}
\theta_{\rm E} = 4\pi\left(\frac{\sigma_v}{c}\right)\frac{D_{ls}}{D_{s}}. 
\end{equation}
%-------------------------------------------------%
Here, $c$ is the speed of light, $\sigma_v$ is the velocity dispersion
of the lens galaxy, and $D_{ls}$ and $D_{s}$ are the respective
angular diameter distance from the source plane to the lens plane and
from the source plane to the observer.

To rotate the lenses with a random orientation angles, we adopt the
transformation below for orientation angle, $\phi$:
%-------------------------------------------------%
\begin{eqnarray}
\label{eqn:rotation}
\begin{bmatrix}

    x_{1}\prime      \\
    x_{2}\prime 
\end{bmatrix}
= 
\begin{bmatrix}
    \cos \phi  &  -\sin \phi      \\
    \sin \phi  &  ~\cos \phi      
\end{bmatrix} 
\begin{bmatrix}
    x_1      \\
    x_2      
\end{bmatrix} .
\end{eqnarray}
%-------------------------------------------------%
From Equations~\ref{eqn:sie}~-~\ref{eqn:rotation}, there are five
independent parameters that uniquely define the lensing system:
$\sigma_v$, ellipticity $q$, orientation angle $\phi$, redshift of the
lens $z_l$, and redshift of the source $z_s$.

% Parameters of lensing system
We choose $\sigma_v$, $q$, and $\phi$ randomly from typical ranges of
a galaxy, where $\sigma_v = [200, 320] km/s$, $q = [0.5, 1.0]$, and
$\phi = [0, 360]$.  The redshift of the lens galaxy is drawn from the
distribution \todo{found in \citet{???}}.  We fix the redshift of the
source to $z_s = 1.0$. We then extract images of 33 bright galaxies
from the CANDELS survey as source galaxies.  We place the source
galaxy images in projected positions close to the caustic of the
lensing system and fix the projected position of the lens galaxy to
the center of the field of view of each image.  Once we have
determined all parameters of the lensing system, we perform
ray-tracing simulations to produce the final lensed image.

%-------------------------------------------------%
\subsubsection{Modeling Lens Galaxies}
%-------------------------------------------------%
We model the light distributions of the lens galaxies with an
elliptical Sersic profile,
%-------------------------------------------------%
\begin{equation}\label{eqn:sersic}
I(R) = I_{\rm eff}~{\rm exp} \left\lbrace -b_{n} \left[ \left(
  \frac{R}{R_{\rm eff}}\right)^{1/n} - 1 \right ] \right\rbrace,
\end{equation}
%-------------------------------------------------%
where, $R = \sqrt[]{x_1^2 /q+x_2^2 q }$, $R_{\rm eff}$ is the
effective radius in units of arcseconds, $I_{\rm eff}$ is the
intensity at the effective radius, $n$ is the index of the Sersic
profile, and $q$ is the ellipticity of the lens galaxy.  We perform a
similar transformation as Equation~\ref{eqn:rotation} to orient the
sources.  We assume that the distribution of light is following that
of mass, so ellipticity and orientation are the same as the SIE model.

We use the COSMOS morphological
catalog \footnote{\url{https://arxiv.org/pdf/1501.04977v2.pdf}} to
match the velocity dispersion with an assigned effective radius,
effective luminosity, and index to the light profile.  We also assume
the light center is on top of the mass center, creating noiseless
images of lens galaxies.

We construct galaxies along the line of sight by cutting light-cones
from the \todo{Hubble Ultra Deep Field\footnote{udf...org}}.  We stack
these images with the lens galaxy image and the lensed source galaxy,
calibrating the magnitude of all three components.

%-------------------------------------------------%
\subsubsection{Producing Mock Images}
%-------------------------------------------------%

The final step is to perform a mock observations.  For HST-like
observations, galaxies along the line of sight already include noise
and a point spread function (PSF).  We therefore stack and magnitude
calibrate all components to produce the final HST-like images, which
are $300\times300~{\rm pix}^2$ with $0.03$~arcsec per pixel.

To produce LSST-like images, we use {\em LensPop} to serially perform
re-binning, convolve with the corresponding PSF, and add
noise\citep{collett_15}.  We created one year LSST-like (LSST1) images
and stacked ten-year LSST-like images (LSST10) to investigate the
performance and usability of our pipeline in future LSST
data. LSST-like images are $45\times45~{\rm pix}^2$ with $0.2$~arcsec
per pixel.

Figure~\ref{fig:mockimages} illustrates sample mock observations with
a strong lensing signature from each telescope.  The left-most column
respectively correspond to a mock HST lensing system with a highly
magnified source galaxy (top) and a less visible image of the source
galaxy (bottom). For the HST-like dataset, many arcs are visually
obvious due to the exquisite spatial resolution.

The middle column corresponds to the same simulated systems, as
observed with over ten years of LSST, LSST10.  The right-most column
corresponds to the simulated systems observed over just one year of
LSST1.  LSST10 images visually exhibit an improved signal to noise
ratio, recovering the arc feature, albeit at a much lower resolution
than with the HST-like image.  The top images of LSST10 and LSST1 show
a visible lensed source galaxy image.  The bottom images are washed
out in the bottom row, where the magnification of the source galaxy is
not as large.  The ground based noise, PSF, and limited resolution of
the LSST1 make visual giant arc identification difficult, except in
systems with the most magnified source galaxies.

@@

Furthermore, we investigate the influence of light from the lens
galaxies on the performance of our lens identification pipeline.  We
generate another set of each HST, LSST1, and LSST10 images without the
lens galaxy.  We label these nHST, nLSST1, and nLSST10.

The final data is then comprised of $6\times10,000$ lens containing
images, and $6\times 10,000$ non-lens containing images.  We also keep
a hold-out set of $6\times 1,000$ lens containing images, and $6\times
1,000$ non-lens containing images.
%-------------------------------------------------%
\subsection{Strong Arc Lensing Identification Pipeline}
%-------------------------------------------------%

@@

%-------------------------------------------------%
\subsection{Feature Extractor: Histogram Oriented Gradients}\label{sec:hog}
%-------------------------------------------------%
% HOG intro - 
Originally created for human detection in computer vision, histogram
oriented gradients (HOG) is a feature extraction method that computes
centered horizontal and vertical gradients. HOG is relatively robust
to noise in the image, and is a fairly fast transform that describes
edges.  Details can be found in \citet{dalalandtriggs_05}, but we
describe the procedure here.  The end result of HOG is a one
dimensional histogram computed as follows.

% Division of the image  
HOG first divides the image into blocks of \wording{50\% overlap}.
Each block contains $m\times m$ cells-per-block that each contain
$n_{pix}\times n_{pix}$ pixels-per-cell.  The computed gradient
orientation is quantized into $N_{orient}$ bins.

@@

%-------------------------------------------------%
\subsection{Machine Learning Algorithm: Logistic Regression}\label{sec:LR}
%-------------------------------------------------%
The problem of detecting gravitational lenses in images falls under
the general category of {\em{classification}} in machine learning. In
general the task is to find a function that assigns data points $x$ to
one of two or more classes, denoted by the class label $y$.  This is
equivalent to specifying a decision boundary, or decision boundaries
between the classes in the space of the data points.  (Compare this to
{\em{regression}} in which the task is to find a function $y=f(x)$,
where $y$ is a continuous, rather than discrete, variable.) In our
case, we have two classes: lens and non-lens containing images, and
the data points $x$ are the HOG feature vectors transformed from the
training set data.  In this paper we use the Logisitic Regression (LR)
algorithm, for which the decision boundary is a hyperplane .  (The
equivalent in the regression setting would be linear regression.) In
LR we determine the optimal hyperplane by minimizing the objective
function
\begin{equation}
\label{eqn:log_reg_objective}
L(A,b) = \sum_i \log \left[1 + \exp \left(-y_i (A \cdot x_i + b) \right) \right]
\end{equation}
where $x_i$ is a data point (HOG feature vector), $y_i$ is the
known label for that data point ($1$ for a lens containing
image, $-1$ for a non-lens containing), and $A$ and $b$ are
the parameters of the hyperplane. Eq. \ref{eqn:log_reg_objective}
is to be minimized with respect to $A$ and $b$.
Other more complicated machine learning algorithms exist which
do not necessarily produce a linear decision boundary, such as
Support Vector Machines (SVM), Random Forests, and Neural Networks.
training set data \citep{hastie_09}.

@@

% Benefits of our method
We emphasize that classical linear classifiers, such as logistic
regression, is scalable and relatively easy to parallelize with open
source tools such as \todo{\em Spark}\footnote{...spark...org}}.
  Initial checks indicated that HOG features are well separated enough
  for a linear classifier algorithm.  Reproducible results do not
  require supercomputing resources.

% Impact of work
These results indicate that effective image classification of
galaxy-galaxy strong lensing systems do not require large scale
computing resources.

% Caveat of the mock data set
One major caveat of our resuls is the fact that our mock data likely
does not describe the full range of lens and non-lens images that will
be observed.  For example, the along the line of sight galaxies in our
training and test images all come from the Hubble Ultra Deep field,
sampling a smaller range of potential systems that are not associated
with a lensed image.  We will address these shortcomings in subsequent
work.

>>>>

% Image identification works: visual, ``robot'' identification -
% ArcFinder, RingFinder..., and citizen science (space warps).  Many
% of these rely on some explicit aspect of the image's morphology
% (give example), filtering, ... ....  Specifics of the methodology
% how the instrument affects the strong lensing signal, and the
% objects of interest.
Over the last decade, infrastructure for both large scale visual and
automated image classification emerged. By nature, the human eye is
one of the best discriminators for image classification.  Visual arc
identification has been effective through the use of citizen science
platforms.  {\em SpaceWarps} is an example of citizen science based
image classification of strong lensing systems in Canada France Hawaii
Telescope Science (CFHTLS) telescope observations
\citep{marshall_etal16,more_etal16}.  These platforms are quite
successful for a dataset like CFHTLS; here, 3,000 candidate images
were identified in eight months, resulting in 89 final candidates.
However, future datasets like Euclid \citep{oguriandmarshall_10} and
LSST expect to find tens of thousands galaxy-scale strong lenses
\citep{pawase_etal12}.  The volume of upcoming data challenges the
scalability of a pure citizen science approach.

@@

% Convolutional and residual
\wording{The most advanced use of neural networks for strong lensing
  classification operate directly in image space.
  \citep{petrillo_etal17} used mock Kilo Degree Survey data to train
  convolutional neural networks, with a training set size of six
  million images. \citet{lanusse_etal17} used state-of-the-art
  residual learning to also work directly in the image space with
  minimal image pre-processing.  A major strength of the
  \citet{lanusse_etal17} implementation is that residual learning has
  been found to be easier to train in deep neural networks, and can
  outperform implementations of convolutional neural networks.}

%  Relevance of our work
\wording{While there has been a recent surge in the use of neural
  networks applied to astronomy image classification problems, these
  implementations are not readily scalable, nor are the necessary
  computational resources and expertise easily accessible to the
  entire scientific community.  We present a supervised classification
  pipeline built with open source tools, tested on mock HST and LSST
  data, which will also be made publicly available.  Our goal is to
  centralize an open source pipeline that is general enough for the
  user to select or add appropriate image processing techniques to
  augment the features of the lensed image, train and test data with
  different machine learning techniques, and to select the sequence of
  methods that provides optimal model performance.  Results are
  reproducible on personal computers, as both the pipeline and the
  data will be publicly distributed.}

% Summary of paper 
This first paper serves as both a presentation of the mock data set
and as a thorough case study for our pipeline with one feature
extraction and classical machine learning method, respectively, the
{\em histogram of oriented gradients} (HOG) and {\em logistic
  regression} (LR).  HOG is a fast feature extraction procedure that
quantifies edges in images, commonly used to identify humans in
security software.  LR is a linear classifier, making its scalability
relatively straightforward with existing open source tools.  This
framework is easily extendable to other feature methods and machine
learning algorithms for supervised classification.

@@

%-------------------------------------------------%
\section{Methodology}
\label{sec:methods}
%-------------------------------------------------%
\subsection{Mock Images}
%-------------------------------------------------%
We use mock {\em Hubble Space
  Telescope}\footnote{https://www.spacetelescope.org/} (HST) and {\em
  Large Synoptic Sky Telescope}\footnote{https://www.lsst.org} (LSST)
images generated with {\em PICS} (Pipeline for Images of Cosmological
Strong lensing) \citep{li_etal16} and {\em LensPop} \citep{collett_15}
for training and testing.

There are 10,000 lens containing mock observations and 10,000
non-lens containing mock observations for running our parameter search. 
We keep a hold-out test set of 1,000 lens containing mock observations
and 1,000 non-lens containing mock observations on which we evaluate
the final trained model.

Mock observations of lensing systems include the lens galaxy, lensed
images of the source galaxy, and galaxies along the line of sight.
Mock observations of non-lensing systems include all but the images of
a lensed source galaxy.  We convolve each of the $2\times 10,000$
train/test images and $2\times 1,000$ hold-out images to produce three
separate sets of mock ``observations''.  These each have equal numbers
of lens and non-lens containing systems: (1) HST-like observations,
(2) LSST-like observations over the span of one year, and (3)
LSST-like observations over the span of ten years.  We respectively
label these observations as HST, LSST1, and LSST10.

%-------------------------------------------------%
\subsubsection{Modeling the Mass Distribution of Lens Galaxies}
%-------------------------------------------------%

To produce simulated lensed galaxies, we first model the mass of lens
galaxies as the Singular Isothermal Ellipsoid (SIE). This model is
both analytically tractable and is consistent with models of
individual lenses and lens statistics on the length scales relevant
for strong lensing
\citep[e.g.]{koopmans_etal06,gavazzi_etal07,dye_etal08,liandchen_09}.

The normalized convergence map of SIE model is defined as:
%-------------------------------------------------%
\begin{equation}\label{eqn:sie}
\kappa = \frac{\theta_{\rm E}}{2}\frac{1}{\sqrt{x_1^2/q+x_2^2 q}},
\end{equation}
%-------------------------------------------------%
where $\theta_{\rm E}$ is the Einstein Radius and $q$ is axis ratio.
The Einstein radius can be calculated from the redshift of the
lens, the redshift of the source, and the velocity dispersion of the
lens galaxy as follows,
%-------------------------------------------------%
\begin{equation}
\theta_{\rm E} = 4\pi\left(\frac{\sigma_v}{c}\right)\frac{D_{ls}}{D_{s}}, 
\end{equation}
%-------------------------------------------------%
Here, $c$ is the speed of light, $\sigma_v$ is the velocity dispersion
of the lens galaxy, $D_{ls}$ and $D_{s}$ are respectively the angular
diameter distances from the source plane to the lens plane and from
the source plane to the observer.

To rotate the lenses with random orientation angle $\phi$, we adopt
the transformation below:
\begin{eqnarray}
\label{eqn:rotation}
\begin{bmatrix}
    x_{1}\prime      \\
    x_{2}\prime 
\end{bmatrix}
= 
\begin{bmatrix}
    \cos \phi  & -\sin \phi      \\
    \sin \phi  & ~\cos \phi      
\end{bmatrix} 
\begin{bmatrix}
    x_1      \\
    x_2      
\end{bmatrix}.
\end{eqnarray}

From Equations~\ref{eqn:sie}~-~\ref{eqn:rotation}, there are five
independent parameters from which we can derive the lensing map:
velocity dispersion $\sigma_v$, axis ratio or ellipticity $q$,
orientation angle $\phi$, redshift of the lens $z_l$, and redshift of
the source $z_s$.

We choose $\sigma_v$, $q$, and $\phi$ randomly from typical ranges of
observed galaxies: $\sigma_v \in [200, 320]~km/s$, $q \in [0.5, 1.0]$,
and $\phi \in [0, 360]$.  We obtain the redshift of the lens galaxy by
matching the velocity dispersion from our simulations to the catalog
of elliptical galaxies in the COSMOS survey from \citet{zahid_etal15},
resulting in lens galaxy redshifts in the range, $z_l \in [0.2,0.7]$.

%-------------------------------------------------%
\subsubsection{Modeling Images of the Lens Galaxies}
%-------------------------------------------------%
We model the light distributions of the lens galaxies with an
elliptical Sersic profile,
%-------------------------------------------------%
\begin{equation}\label{eqn:sersic} 
I(R) = I_{\rm eff}~{\rm exp} \left\lbrace -b_{n} \left[ \left(
  \frac{R}{R_{\rm eff}}\right)^{1/n} - 1 \right ]
\right\rbrace 
\end{equation}
%-------------------------------------------------%
where, $R = \sqrt[]{x_1^2 /q+x_2^2 q }$, $R_{\rm eff}$ is the
effective radius in units of arcseconds, $I_{\rm eff}$ is the
intensity at the effective radius, $n$ is the index of the Sersic
profile, and $q$ is the ellipticity of the lens galaxy.  We perform a
similar transformation as Equation~\ref{eqn:rotation} to orient the
source galaxies and assume that the distribution of light follows that
of mass.  The ellipticity and orientation are therefore the same as in
the SIE model.

We use the COSMOS morphological catalog \citep{zahid_etal15} to match
the velocity dispersion with an assigned effective radius, effective
luminosity, and index to the light profile.  We also assume the light
center is on top of the mass center, creating noiseless images of lens
galaxies.

We construct galaxies along the line of sight by cutting light-cones
from the Hubble Ultra Deep
Field\footnote{http://www.spacetelescope.org/science/deep\_fields/}.
We stack these images with the lens galaxy image and the lensed source
galaxy, calibrating the magnitude of all three components.

%-------------------------------------------------%
\subsubsection{Modeling Images of the Source Galaxies}
%-------------------------------------------------%
The background source galaxies come from a set of detailed images of
low-redshift bright galaxies ($z\sim0.45$) that have been extracted
from mosaics produced by the CANDELS team
\citep{grogin_etal11,koekemoer_etal11} and selected from the CANDELS
UDS catalog \citep{galametz_etal13}.  We rescale these background
source galaxies, set the source redshift to $z_s=0.2$, and select
source positions near caustics of the lensing system.  We then fix the
projected position of the lens galaxy at the center of the field of
view for each image.

We perform ray-tracing simulations using the modeled galaxy images and
parameters.  The lens galaxy, source galaxy, and parameters of the
image simulations (i.e., $300 \times 300~{\rm pixels}^2$ with
0.03~arcsec/pixel for an HST-like image) are inputs to the simulation
that produce ideal lensed images with the appropriate resolution.

%-------------------------------------------------%
\subsubsection{Mock Observing}
%-------------------------------------------------%

The final part is to perform the mock observation from the ideal
images produced in the previous step. For HST-like observations,
galaxies along the line of sight already include noise and a point
spread function (PSF).  We therefore stack and magnitude calibrate all
components to produce the final HST-like images, which are
$300\times300~{\rm pix}^2$ with $0.03$~arcsec per pixel.

For LSST-like observations, we use the {\em LensPop} software
\citep{collett_15}.  We resample images to match the detector pixel
scale and convolve the resampled image with a circularly symmetric
Gaussian Point Spread Function discretised at the same pixel scale. To
generate a noisy realization of the image, we assume a Poisson model
based on the sky plus signal, and an additional Gaussian read-out
noise. Parameters for these simulations follow \citet{collett_15} and
are based on the LSST observation simulator \citep{connolly_etal10}.

To account for variations in seeing and sky-brightness over the course
of the survey, we draw each simulated exposure from a stochastic
distribution of these parameters. We then consider two different
strategies to use the simulated exposures. First, we build one
single-epoch image for each field (hereafter labeled as LSST1) by
keeping only the best seeing exposure.  Second, we build another
``worst-case'' stacked image by degrading all individual exposures to
match the one with the worst seeing and co-add all exposures to a
single image (hereafter labeled as LSST10).  These two sets of images
will allow us to investigate the trade-off between resolution and
signal-to-noise for our automated lens search.

Figure~\ref{fig:mockimages} illustrates sample mock observations with
a strong lensing signature from each telescope.  The left-most column
corresponds to a mock HST lensing system with a highly magnified
source galaxy (top) and a less visible image of the source galaxy
(bottom). For the HST-like dataset, many arcs are visually obvious due
to the exquisite spatial resolution and quality of space based
imaging.

The middle column corresponds to the same simulated systems for
LSST10.  The right-most column corresponds to the simulated systems of
LSST1.  These images have resolution $45\times45~{\rm pixels}^2$ with
$0.18$~arcsec/pixel.  LSST10 images visually exhibit the improved
signal to noise ratio, recovering the arc feature, albeit at a much
lower resolution than with the HST-like image or the LSST1 image.  The
top images of LSST10 and LSST1 show a visible lensed source galaxy
image.  The bottom images are washed out in the bottom row, where the
magnification of the source galaxy is not as large.  The ground based
noise, PSF, and limited resolution of the LSST1 make visual giant arc
identification difficult, except in systems with the most magnified
source galaxies.

@@

Furthermore, we investigate the influence of light from the lens
galaxies on the performance of our lens identification pipeline.  We
generate another set of each HST, LSST1, and LSST10 images without the
lens galaxy.  We respectively label these nHST, nLSST1, and nLSST10.

The final data set is then comprised of $6\times10,000$ lens
containing images, and $6\times 10,000$ non-lens containing images.
We also keep a hold-out set of $6\times 1,000$ lens containing images,
and $6\times 1,000$ non-lens containing images.

%-------------------------------------------------%
\subsection{Strong Arc Lensing Identification Pipeline}
%-------------------------------------------------%

@@

%-------------------------------------------------%
\subsubsection{Feature Extractor: Histogram Oriented Gradients}\label{sec:hog}
%-------------------------------------------------%
% HOG intro - 
Originally created for human detection in computer vision, histogram
oriented gradients (HOG) is a feature extraction method that computes
centered horizontal and vertical gradients. HOG is relatively robust
to noise in the image, and is a fairly fast transform that describes
edges.  Details can be found in \citet{dalalandtriggs_05}, but we
describe the procedure here.  The end result of HOG is a one
dimensional histogram computed as follows.

% Division of the image  
HOG first divides the image into blocks of 50\% overlap.  Each block
contains $m\times m$ cells-per-block that each contain $n_{pix}\times
n_{pix}$ pixels-per-cell.  The computed gradient orientation is
quantized into $N_{orient}$ bins.

@@

%-------------------------------------------------%
\subsubsection{Machine Learning Algorithm: Logistic Regression}\label{sec:LR}
%-------------------------------------------------%
The problem of detecting gravitational lenses in images falls under
the general category of {\em{classification}} in machine learning. In
general the task is to find a function that assigns data points $x$ to
one of two or more classes, denoted by the class label $y$.  This is
equivalent to specifying a decision boundary, or decision boundaries
between the classes in the space of the data points.  (Compare this to
{\em{regression}} in which the task is to find a function $y=f(x)$,
where $y$ is a continuous, rather than discrete, variable.) In our
case, we have two classes: lens and non-lens containing images, and
the data points $x$ are the HOG feature vectors transformed from the
training set data.  In this paper we use the Logisitic Regression (LR)
algorithm, for which the decision boundary is a hyperplane .  (The
equivalent in the regression setting would be linear regression.) In
LR we determine the optimal hyperplane by minimizing the objective
function
\begin{equation}
\label{eqn:log_reg_objective}
L(A,b) = \sum_i \log \left[1 + \exp \left(-y_i (A \cdot x_i + b) \right) \right]
\end{equation}
where $x_i$ is a data point (HOG feature vector), $y_i$ is the
known label for that data point ($1$ for a lens containing
image, $-1$ for a non-lens containing), and $A$ and $b$ are
the parameters of the hyperplane. Eq. \ref{eqn:log_reg_objective}
is to be minimized with respect to $A$ and $b$.
Other more complicated machine learning algorithms exist which
do not necessarily produce a linear decision boundary, such as
Support Vector Machines (SVM), Random Forests, and Neural Networks.
training set data \citep{hastie_09}.

@@

% Benefits of our method
\wording{We emphasize that classical linear classifiers, such as
  logistic regression, is scalable and relatively easy to parallelize
  with open source tools such as {\em
    Spark}\footnote{http://spark.apache.org/}.  Initial checks
  indicated that HOG features are well separated enough for a linear
  classifier algorithm.  Finally, our methods enable reproducible
  results that do not require supercomputing resources.

% Impact of work
These results indicate that effective image classification of
galaxy-galaxy strong lensing systems do not require large scale
computing resources or expertise.}

% Caveat of the mock data set
One major caveat of our resuls is the fact that our mock data likely
does not describe the full range of lens and non-lens images that will
be observed.  For example, the along the line of sight galaxies in our
training and test images all come from the Hubble Ultra Deep field,
sampling a smaller range of potential contaminants that are not
associated with a lensed image.  Also, the redshift of the source
galaxies have been fixed to $z_s=.2$.  We will address these
shortcomings in subsequent work.
